---
title: "Tampa Bay Area 2016 Fatal Crashes Analysis"
output:
  html_document
---
#### _By Sharon Tong, University of South Florida - Tampa_

[View on GitHub](https://github.com/SWTong/fatal_crashes)

```{r include = FALSE}
library("leaflet")
library("magrittr") 
library("dplyr")

#get directory path of data set
setwd("C:\\Users\\Sharon\\Desktop\\fatal_crashes")

#read the tbay csv file
tbay_acc <- read.csv("tbay_acc.csv")
hb <- filter(tbay_acc, tbay_acc$COUNTY_NAME == "HILLSBOROUGH")
```

``` {r map_label, include = FALSE}

popupInfo <- paste("<b>Month: </b>", tbay_acc$MONTH_NAME,
                 "<br>", "<b> Day of Week: </b>", tbay_acc$DAY_WEEK_NAME,
                 "<br>", "<b> Trafficway1: </b>", tbay_acc$TWAY_ID,
                 "<br>", "<b> Trafficway2: </b>", tbay_acc$TWAY_ID2,
                "<br>", "<b> Intersection Type: </b>", tbay_acc$TYP_INT_NAME,
                 "<br>", "<b> County: </b>", tbay_acc$COUNTY_NAME,
                 "<br>", "<b> City: </b>", tbay_acc$CITY_NAME
                  )
```

```{r tpabay_map, include=FALSE}
tbay_map <- leaflet(data=tbay_acc, width = "100%") %>% 
addTiles() %>% 
addMarkers(data = tbay_acc, lng = ~ LONGITUD, lat = ~ LATITUDE, 
                  popup = popupInfo,
                   clusterOptions = markerClusterOptions())
```

```{r,echo=FALSE}
tbay_map
```

###The Problem
Lots of accidents happen all over the country. According to %
Florida is ranked % in the year % in terms of the number of accidents that occurred. A lot of research has been done on the demographics of accidents already, such as age and gender, so I will be mainly focusing on the characteristics of the accidents themselves such as month, day of week, and location, particularly in the Tampa Bay Area.


###The Data
The data I will be using will come from one of the csv files named `accidents.csv` that contains data about __fatal__ accidents (accidents where at least one death occurred) that occurred in all U.S. states in 2016. The csv file can be obtained from within a zipped folder from the U.S. Fatality Analysis Reporting System (FARS) [here](ftp://ftp.nhtsa.dot.gov/fars/2016/National/). After downloading the file, I obtained additional data from the U.S. Geographic Locator Codes site [here](https://www.gsa.gov/reference/geographic-locator-codes/glcs-for-the-us-and-us-territories) to decode some of the geographic codes in the original accidents data set. I Extracted, Transformed, and Loaded the data into Microsoft SQL Server to store the data and organized the data into a data warehouse to prepare for analysis. 

The resulting clean data set that R will be analyzing is [this](tbay_acc.csv).

Here is a preview of the data set:

```{r table}
#Preview first 50 rows of the data set, list 10 entries per page
library(DT)

preview <- tbay_acc[1:50,]
table <- datatable(preview, options = list(pageLength = 10,scrollX='400px'))
```
```{r, echo=FALSE}
table
```

###Descriptive Statistics

Let's take a look at the structure of the data set by running the R code below. The result shows that there are 705 observations (rows) and 25 variables (columns). On the left-hand side, R displays a list of all the columns and on the right-hand side, their data types and examples of some of their values. "Int" represents integer data, "Factor" represents qualitative data, and "num" represents non-integer data.


```{r tpabay}
str(tbay_acc)
```

__Question: How many fatal accidents occurred in each county in the Tampa Bay Area? Is there a difference in the number of accidents between rural and urban areas?__

To answer this question, we can create a bar chart based on our data that includes the county name, rural and urban area types, and the frequency of each accident.

```{r rururb, include = FALSE}
#subset tbay csv data set to only show accidents in rural areas
tbay_rur <- filter(tbay_acc, tbay_acc$RUR_URB_NAME == "Rural")

#subset tbay csv data set to only show accidents in urban areas
tbay_urb <- filter(tbay_acc, tbay_acc$RUR_URB_NAME == "Urban")

#create a new data set that only has county name, rural, and count of accidents to prepare for plotly bar chart
ctyrur <- data.frame(tbay_rur %>% group_by(tbay_rur$COUNTY_NAME,tbay_rur$RUR_URB_NAME) %>% tally())

#rename ctyrur column n to be called Rural to prepare for plotly bar chart
ctyrur <- rename(ctyrur, "Rural" = "n", "County" = "tbay_rur.COUNTY_NAME")
  
#remove the ctyrur column tbay_rur.RUR_URB_NAME to prepare for plotly bar chart
ctyrur <- select(ctyrur,-"tbay_rur.RUR_URB_NAME")

#create a new data set that only has county name, urban, and count of accidents to prepare for plotly bar chart

ctyurb <- data.frame(tbay_urb %>% group_by(tbay_urb$COUNTY_NAME,tbay_urb$RUR_URB_NAME) %>% tally())

ctyurb <- rename(ctyurb, "Urban" = "n", "County" = "tbay_urb.COUNTY_NAME")

ctyurb <- select(ctyurb, -"tbay_urb.RUR_URB_NAME")

ctyrururb <- bind_cols(ctyrur, ctyurb)

#merge rural and urban data sets to prepare for plotting
ctyrururb <- left_join(ctyrur, ctyurb, by=c("County"))

```

```{r rururb_barchart, message = FALSE}

#plot a bar chart using the Plotly library for merged rural and urban data set created from tbay_acc.csv
library("plotly")

rururb_bar <- plot_ly(ctyrururb, x = ~ctyrururb$County, y = ~ctyrururb$Rural, type = 'bar', name = 'Rural', marker = list(color = 'rgb(49,130,189)')) %>%
  add_trace(y = ~ctyrururb$Urban, name = 'Urban', marker = list(color = 'rgb(204,204,204)')) %>%
  layout(title = "Tampa Bay Fatal Accidents by County and Rural/Urban Area in 2016",
        xaxis = list(title = "County", tickangle = -45),
         yaxis = list(title = "Number of Accidents"),
         margin = list(b = 100),
         barmode = 'group')
rururb_bar
```

From the bar chart, we can see that Hillsborough County had the highest number of fatal crashes in 2016 compared to other Tampa Bay counties. There appears to be more fatal accidents occurring in urban areas compared to rural areas. 

Note: this chart does not include areas that are not rural or urban even though non-rural and non-urban data is available within the tbay_acc.csv file.

__Question: Where do most fatal accidents occur in Hillsborough County?__

To answer this question, we can create a map based on rows from our tbay_acc.csv where the county name is Hillsborough
and apply a cluster accidents that are within a short distance from each other to make the map more readable.

```{r hb_map_code}
popupInfo2 <- paste("<b>Month: </b>", hb$MONTH_NAME,
                 "<br>", "<b> Day of Week: </b>", hb$DAY_WEEK_NAME,
                 "<br>", "<b> Trafficway1: </b>", hb$TWAY_ID,
                 "<br>", "<b> Trafficway2: </b>", hb$TWAY_ID2,
                "<br>", "<b> Intersection Type: </b>", hb$TYP_INT_NAME,
                 "<br>", "<b> County: </b>", hb$COUNTY_NAME,
                 "<br>", "<b> City: </b>", hb$CITY_NAME
                  )

hb_map <- leaflet(data=hb) %>% 
addTiles() %>% 
addMarkers(data = hb, lng = ~ LONGITUD, lat = ~ LATITUDE, 
                  popup = popupInfo2,
                   clusterOptions = markerClusterOptions())
```

```{r, echo=FALSE}

hb_map
```

From the map of Hillsborough County, we can see that Tampa had the highest number of fatal accidents, leading with about 57 accidents. Within Tampa,  I-275 was the location where most fatal accidents occurred.

__Question: What months had the highest number of fatal accident in the Tampa Bay Area in 2016?__

To answer this question, we can create another bar chart based on our data that includes the month and the frequency of each accident.

```{r, month_bar}
#create a data set with month and number of accidents from the tbay_acc csv file, 
#rename the month data set columns, and 
#plot the month data set using the Plotly library
library("plotly")

months_set <- data.frame(tbay_acc %>% group_by(tbay_acc$MONTH_NAME) %>% tally())
months_set <- rename(months_set, "Month" = "tbay_acc.MONTH_NAME", "AccCount" = "n")
month_plot <- plot_ly(months_set, x = ~months_set$Month, y = ~months_set$AccCount, type = 'bar') %>%
  layout(title="Tampa Bay Fatal Accidents by Month in 2016",
       xaxis = list(title = "Month", tickangle = -45),
         yaxis = list(title = "Number of Accidents")
       )
month_plot
```

From the bar chart, we can see that the month with the highest number of fatal accidents in 2016 in the Tampa Bay Area is May with 69 accidents, followed by August with 65 accidents.

###References

1. [Using R for Crime Analysis](https://wetlands.io/maps/Crime-Analysis-Using-R.html#aggregate_data) by Dr. Qiusheng Wu, State University of New York at Binghamton
2. [Bar Charts in R](https://plot.ly/r/bar-charts/) by Plotly
3. [Create a website with R](https://www.youtube.com/watch?v=QYr06j_JMdo) by Alex Shpenev
4. [Leaflet for R - Markers](https://rstudio.github.io/leaflet/markers.html) by RStudio